---
title: "Hold'em Strategies"
author: "Alex"
date: "2023-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(tidyr)
library(dplyr)
options(dplyr.summarise.inform = FALSE)
library(scales)
library(ggplot2)
library(knitr)
```

# Overview

In Texas Hold'em, two cards, known as hole cards, are dealt face down to each player, and then five community cards are dealt face up in three stages. Each player seeks the best five card poker hand from any combination of the seven cards; the five community cards and their two hole cards. This document is a game simulation on Texas Hold'em in a heads up(1v1) scenario. The goal of this simulation to test the better strategy between only playing if player has a pair or two cards of the same suit and when player has two cards that is less than 4 cards apart.

# Setup

The simulation setup creates a deck of 52 cards for each simulated hand. 5 community cards, 2player cards and 2 enemy cards are dealt from each deck for every hand. This cards are wrangled into a data frame that includes the card value, player and hand. Showing the first 2 hands.

```{r}
no_of_hands= 10000
deck <- data.frame(
  suit = rep(c("Diamonds", "Clubs", "Hearts", "Spades"), each =13),
  number = rep(c("ACE",2:10,"JACK","QUEEN","KING"), times =4),
  value = rep(c(14,2:13), each = 4)
)

players = c("Me","Me","Enemy","Enemy","Community","Community","Community","Community","Community")

for (x in 1:no_of_hands) {
  if (x == 1){
    Game <- data.frame(
  hand = rep(x,each = 9),
  player = players,
  sample_n(deck, 9,replace = FALSE))
  }
  else{
    temp <- data.frame(
  hand = rep(x,each = 9),
  player = players,
  sample_n(deck, 9,replace = FALSE))
  Game <-bind_rows(Game,temp)
  }}
head(Game,18)
```

# Checking the winner

Here are the ranking of combinations to determine the winner. (From Strongest to weakest)

1. Royal flush
A, K, Q, J, 10, all the same suit.

2. Straight flush
Five cards in a sequence, all in the same suit.

3. Four of a kind
All four cards of the same number.

4. Full house
Three of a kind with a pair.

5. Flush
Any five cards of the same suit, but not in a sequence.

6. Straight
Five cards in a sequence, but not of the same suit.

7. Three of a kind
Three cards of the same number.

8. Two pair
Two different pairs.

9. Pair
Two cards of the same number.

10. High Card
When you haven't made any of the hands above, the highest card wins.

Creating a function to determine the winner.

```{r}
# Calculate the value of a hand
hand_value <- function(hand) {
  # Create a subset of the hand with only the community cards
  community_cards <- hand[hand$player == "Community", ]
  # Sort the community cards by their value
  community_cards <- community_cards[order(community_cards$value), ]
  # Combine the two cards in each player's hand
  player_hands <<- split(hand[hand$player != "Community", ], hand[hand$player != "Community", ]$player)
  player_hands <- lapply(player_hands, function(x) x[order(x$value), ][1:2, ])
  player_hands_community <- lapply(player_hands, function(x) rbind(x,community_cards))
  # Check if each player has a royal flush
  has_royal_flush <<- sapply(player_hands_community, function(x) {
    royal <- (all(x$number %in% c("ACE", "KING", "QUEEN", "JACK", 10)))
              
    count <- x %>% group_by(hand,suit) %>% 
  summarize(matches = n()) %>% mutate(flush_me = matches >= 5) %>% filter(flush_me == TRUE) %>% nrow()
    
    if (royal == TRUE & count>0) return(1) else return(0)
  })
  
  # Check if each player has a straight flush
  has_straight_flush <<- sapply(player_hands_community, function(x) {
    if (any(x$number %in% c("ACE"))){
      tempace <- x[x$number %in% c("ACE"),]
      tempx <- x
      rbind(tempx, tempace)
      scount <- tempx %>% arrange(hand,number) %>% group_by(hand, value) %>%
  summarize(matches = n()) %>% mutate(straight = lag(value,4, default = -10)== value-4) %>% filter(straight == TRUE)%>% nrow()
    }
    else{
      scount <- x %>% arrange(hand,number) %>% group_by(hand, value) %>%
  summarize(matches = n()) %>% mutate(straight = lag(value,4, default = -10)== value-4) %>% filter(straight == TRUE)%>% nrow()
    }
    
    
    fcount <- x %>% group_by(hand,suit) %>% 
  summarize(matches = n()) %>% mutate(flush_me = matches >= 5) %>% filter(flush_me == TRUE) %>% nrow()
    
    if (scount>0 & fcount>0) return(1) else return(0)
  })
  
  # Check if each player has four of a kind
  has_four_of_a_kind <<- sapply(player_hands_community, function(x) {
    count <- x %>% group_by(hand,number) %>% summarize(matches = n())%>% mutate(four_of_a_kind = matches == 4) %>% filter(four_of_a_kind == TRUE) %>% nrow() 
    
    if (count >0) return(1) else return(0)
  })
  
  # Check if each player has a full house
  has_full_house <<- sapply(player_hands_community, function(x) {
    df_a <<- x %>% group_by(hand, value) %>% summarize(counts = n()) %>% filter(counts == 3)
    if (length(df_a$value) == 0) return (0) else{

    df_a <- df_a %>% filter(value == max(value))
    df_b <- x %>% filter(value != df_a$value) %>% group_by(hand, value) %>% summarize(counts = n()) %>% filter(counts >= 2)
    
    is_fullhouse <- (length(df_a$value) + length(df_b$value) >= 2)
    
    if (is_fullhouse == TRUE) return(1) else return(0)
  }})
  
  # Check if each player has a flush
  has_flush <<- sapply(player_hands_community, function(x) {
    count <- x %>% group_by(hand,suit) %>% 
  summarize(matches = n()) %>% mutate(flush_me = matches >= 5) %>% filter(flush_me == TRUE) %>% nrow()
    
    if (count>0) return(1) else return(0)
  })
  
  # Check if each player has a straight
  has_straight <<- sapply(player_hands_community, function(x) {
    if (any(x$number %in% c("ACE"))){
      tempace <- x[x$number %in% c("ACE"),]
      tempx <- x
      rbind(tempx, tempace)
      count <- tempx %>% arrange(hand,number) %>% group_by(hand, value) %>%
  summarize(matches = n()) %>% mutate(straight = lag(value,4, default = -10)== value-4) %>% filter(straight == TRUE)%>% nrow()
    }
    else{
      count <- x %>% arrange(hand,number) %>% group_by(hand, value) %>%
  summarize(matches = n()) %>% mutate(straight = lag(value,4, default = -10)== value-4) %>% filter(straight == TRUE)%>% nrow()
    }
    
    
    if (count>0)  return(1) else return(0)
  })
  
  #Check if each player has three of a kind
  has_three_of_a_kind <<- sapply(player_hands_community, function(x) {
    count <- x %>% group_by(hand,number) %>% summarize(matches = n())%>% mutate(three_of_a_kind = matches == 3) %>% filter(three_of_a_kind == TRUE) %>% nrow() 
    
    if (count >0) return(1) else return(0)
  })
  
  #Check if each player has two pairs
  has_two_pairs <<- sapply(player_hands_community, function(x) {
    count <- x %>% group_by(hand,number) %>% summarize(matches = n())%>%
  filter(matches >= 2) %>% group_by(hand) %>% summarize(matches_two = n()) %>%  mutate(two_pair = matches_two >= 2) %>%
  filter(two_pair == TRUE) %>% nrow()
    
    if (count>0) return(1) else return(0)
  })
  
  #Check if each player has one pair
  has_one_pair <<- sapply(player_hands_community, function(x) {
    count <- x %>% group_by(hand,number) %>% summarize(matches = n())%>% mutate(pair = matches >= 2) %>% filter(pair == TRUE) %>% nrow() 
    
    if (count >0) return(1) else return(0)
  })
  
  #Calculate high card of each hand
  hand_strength <<- sapply(player_hands, function(x) {
    max(x$value)
  })
  
  #Determine the winner
  winner <- ifelse(sum(has_royal_flush) > 0, 
                   names(has_royal_flush[has_royal_flush == 1]), 
            ifelse(sum(has_straight_flush) > 0,
                   ifelse(sum(has_straight_flush) > 1,           names(which.max(hand_strength)),
                          names(has_straight_flush[has_straight_flush == 1])),
            ifelse(sum(has_four_of_a_kind) > 0,
                   ifelse(sum(has_four_of_a_kind) > 1,           names(which.max(hand_strength)),
                          names(has_four_of_a_kind[has_four_of_a_kind == 1])),
            ifelse(sum(has_full_house) > 0, 
                   ifelse(sum(has_full_house) > 1,           names(which.max(hand_strength)),
                          names(has_full_house[has_full_house == 1])), 
            ifelse(sum(has_flush) > 0, 
                   ifelse(sum(has_flush) > 1,           
names(which.max(hand_strength)),
                          names(has_flush[has_flush == 1])), 
            ifelse(sum(has_straight) > 0, 
                   ifelse(sum(has_straight) > 1,           names(which.max(hand_strength)),
                          names(has_straight[has_straight == 1])), 
            ifelse(sum(has_three_of_a_kind) > 0,
                   ifelse(sum(has_three_of_a_kind) > 1,           names(which.max(hand_strength)),
                          names(has_three_of_a_kind[has_three_of_a_kind == 1])), 
            ifelse(sum(has_two_pairs) > 0, 
                   ifelse(sum(has_two_pairs) > 1,           names(which.max(hand_strength)),
                          names(has_two_pairs[has_two_pairs == 1])), 
            ifelse(sum(has_one_pair) > 0, 
                   ifelse(sum(has_one_pair) > 1,           names(which.max(hand_strength)),
                          names(has_one_pair[has_one_pair == 1])),
            names(which.max(hand_strength)))))))))))

return(winner)
}

firsthandwinner <- hand_value(Game[Game$hand==1,])
testwinner <- hand_value(Game[Game$hand==15,])

#doing it for all hands
for (i in 1:no_of_hands){
  Game$winner[Game$hand==i]<- hand_value(Game[Game$hand==i,])
}


winners <- by(Game, Game$hand, hand_value)
winnersdf <- data.frame(winner = table(winners))
winningpercentage <- percent(winnersdf$winner.Freq[winnersdf$winner.winners=="Me"]/no_of_hands)
table(winners)
```

Winner of the first hand is `r firsthandwinner`.
Total winning percentage  is `r winningpercentage`.


# Strategy1
Only go when two cards are the same suit or same number

```{r}
strategy1 <- function(Game) {
  
  winnerdf <- Game %>% filter(player== "Me"& winner == "Me") %>% group_by(hand,number) %>% summarize(matches = n())%>% mutate(pair = matches >= 2) %>% filter(pair == TRUE)%>% distinct()
  
  pairwins <- nrow(winnerdf)
  
  winnerdf <- Game %>% filter(player== "Me"& winner == "Me") %>% group_by(hand,suit) %>% summarize(matches = n())%>% mutate(samesuit = matches >= 2)%>% filter(samesuit == TRUE)%>% distinct()
  
  samesuitwins <- nrow(winnerdf)
  
  totalwin <<- Game %>% group_by(hand)%>% filter(player== "Me"& winner == "Me") %>% select(player,winner) %>% distinct() %>%nrow()
  
  strategy1percentage <<- percent((pairwins+samesuitwins)/totalwin)
  
  return(strategy1percentage)
}

strategy1(Game)
```

Winning percentage with strategy1 is `r strategy1percentage` out of total wins.

# Strategy2
Only go when the difference between the cards is less than or equals to 4.

```{r}
strategy2 <- function(Game) {
  
  #TODO Ace
  winnerdf <- Game %>% filter(player== "Me" & winner == "Me")%>% arrange(hand,value) %>% group_by(hand, value) %>%
  summarize(matches = n()) %>% mutate(previous = lag(value,1, default = -10)) %>% mutate(difference = (value - lag(value,1, default = -10)) <=4)%>% filter(value == max(value)) %>% filter(difference == TRUE) %>% select(hand,difference)%>% distinct()
  
  littledifferencewins <- nrow(winnerdf)
  
  totalwin <<- Game %>% group_by(hand)%>% filter(player== "Me"& winner == "Me") %>% select(player,winner) %>% distinct() %>%nrow()
  
  strategy2percentage <<- percent(littledifferencewins/totalwin)
  
  return(strategy2percentage)
}

strategy2(Game)
```

Winning percentage with strategy2 is `r strategy2percentage` out of total wins.

# Conclusion

```{r}
strategydf <- data.frame(strategy = c("Strategy 1","Strategy 2"), winningpercentage = c(strategy1percentage,strategy2percentage))




ggplot(strategydf, aes(x = strategy, y = winningpercentage)) + geom_bar(stat="identity")
```


The percentage of winning for `r strategydf$strategy[strategydf$winningpercentage == max(strategydf$winningpercentage)]` is higher with `r max(strategydf$winningpercentage)` chance of all the winning hands.




